# Annual maximums {#amax}

## Introduction

In this document I will show how to use `floodnetRfa` to perform at-site 
frequency analysis using annual maximum discharge. 
First we will extract the annual maximums from the
daily discharge data of the Saint-John River at Fort Kent (NB).
Note that incomplet year with missing observations will be removed.
In total 88 annual maximums are extracted and showed in a histogram.

```{r, fig.height=5, fig.width=5}
library(floodnetRfa)

anData <- ExtractAmax(flow ~ date, flowStJohn, tol = 365)

npDens <- density(anData$flow)

hist(anData$flow, 
     freq = FALSE, 
     main = '', 
     xlab = 'Discharge',
     ylim = c(0, max(npDens$y)))

lines(npDens, col = 2, lwd = 2)

```

In stationary situation, the annual maximums are assumed to be independant and 
identically distributed.
Risk is then measure in terms of return period that characterizes the 
average time separating two events of the same magnitude.
In practice this is equivalent to calculate the probability $p = 1-1/T$ of a 
fitted distribution.
The test of Mann-Kendall is frequently performed to verify if the data contains 
a significant trend that would invalidate the assumption of stationarity. 
The present data have a p-value of 0.21, which does not suggest the present of 
a trend.

```{r amax-mkendall, fig.height=5, fig.width=7, fig.cap='Trend in the annual maximums'}

plot(flow~date, anData)
MKendall(anData$flow)
```

## Estimation of the flood quantiles

According to extreme value theory, as the number of annual maximums increase, 
their distribution converge to Generalized Extreme Value (GEV) distribution

\begin{equation}
F(x) = \exp\left\{ - \left[ 1 - \kappa \left(\frac{x-\xi}{\alpha} \right) \right]^{1/\kappa} \right\}
(\#eq:amax-gevcdf)
\end{equation}

The GEV distribution in Equation \@ref(eq:amax-gevcdf) can be fitted using the `FitAmax` function. 
The example below shows how the parameter are estimated using the
maximum likelihood method. 
See for instance [@coles_introduction_2001].
A brief summary of the fitted model is reported, including the estimated parameter,
their standard deviation and the sample L-moments.

```{r}
fit <- FitAmax(anData$flow, 'gev', method = 'mle')

print(fit)
```

The flood quantile of the GEV distribution can be obtained from the formula 

$$
x_T = \mu + \frac{\alpha}{\xi}\left[ 1+\log(1/T)^\kappa \right] .
$$
These predicted value are computed using the `predict` function.
In the example below show the flood quantile for return period 2, 5, 10, 20, 50 
and 100. 
The standard deviation of the flood quantiles is estimated using the Delta 
method that assume that the estimated parameter follow a Normal distribution. 

```{r}
yhat <- predict(fit, se = TRUE, ci = 'delta')
```

```{r amax-yhat, echo = FALSE}
knitr::kable(yhat, caption = 'Predicted return levels')
```


## Verification of the model

The return level plot in \@ref(fig:amax-gevreturnlevel) provide a visual 
assessment of the fitted distribution by comparing the sample and 
the predicted flood quantiles.
The graphic below shows a good agreement between the two. 

```{r amax-gevreturnlevel, fig.height= 5,fig.width=7,  fig.cap = 'Return level plot for Saint-John River at Fort Kent, NB.'}
plot(fit, ci = TRUE)
```

Another diagnostic to ensure that the GEV distribution is appropriate is the
Anderson-Darling test. The p-value superior to 0.05 indicates that the 
hypothesis of a GEV cannot be rejected.

```{r important-computing, cache=TRUE}
## Time consuming when estimated by 'mle'
GofTest(fit, nsim = 500)
```





